---
title: "10_MLPS_R_instance_based_learning"
author: "Zhe Zhang (TA - Heinz CMU PhD)"
date: "3/04/2017"
output: 
  html_document:
    css: '~/Dropbox/avenir-white.css'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F, message = F)
```

## Lecture 10: Instance Based Learning

Key specific tasks we covered in this lecture:

* KNN classifier, with different k
* More advanced KNN: different distance, kernel trick/transformation KNN
* kernel density local regression
* local linear regression
* locally weighted polynomial regression

### KNN

```{r}
library(tidyverse)
library(class)

df <- airquality %>% 
  mutate(high_temp = ifelse(Temp > 75, "High", "Low")) %>%
  drop_na() %>%
  select(-Month, -Day) %>%
  mutate_if(.funs = scale, .predicate = is.numeric)

# important to scale!

train_sample <- sample(seq(3), nrow(df), r = T)
train <- df[train_sample != 3, ]
test <- df[train_sample == 3, ]

knn_preds <- knn(train %>% select(-high_temp), 
                 test%>% select(-high_temp), 
                 cl = train$high_temp,
                 k = 4)

# measure performance
table(knn_preds, test$high_temp)
mean(knn_preds == test$high_temp)

# leave 1 out knn.cv
# k = 2
cv_preds <- knn.cv(df %>% select(-high_temp), 
                   cl = df$high_temp, k = 2)
mean(cv_preds == df$high_temp)
# k = 3
cv_preds <- knn.cv(df %>% select(-high_temp), 
                   cl = df$high_temp, k = 3)
mean(cv_preds == df$high_temp)
# k = 4
cv_preds <- knn.cv(df %>% select(-high_temp), 
                   cl = df$high_temp, k = 4)
mean(cv_preds == df$high_temp)
# k = 5
cv_preds <- knn.cv(df %>% select(-high_temp), 
                   cl = df$high_temp, k = 5)
mean(cv_preds == df$high_temp)

# multi-variable knn
df <- iris %>%
    mutate_if(.funs = scale, .predicate = is.numeric)
# important to scale!

train_sample <- sample(seq(3), nrow(iris), r = T)
train <- df[train_sample != 3, ]
test <- df[train_sample == 3, ]

knn_preds <- knn(train %>% select(-Species), 
                 test%>% select(-Species), 
                 cl = train$Species,
                 k = 4)
table(knn_preds, test$Species)
mean(knn_preds == test$Species)
```


### More Advanced KNN

See the `kknn` package and the `kknn` command. It has options for using a kernel distance options, and customizing the Minkowski distance. See this [manual](https://cran.r-project.org/web/packages/kknn/kknn.pdf). 

In general, using your own distance function may be possible with the `FastKNN` package? See this [StackOverflow question](http://stackoverflow.com/questions/23449726/find-k-nearest-neighbors-starting-from-a-distance-matrix)


### Kernel Density Regression

Predict new points from kernel density using the `x.points` option in `ksmooth`.

```{r}
df <- airquality %>% 
  mutate(high_temp = ifelse(Temp > 75, "High", "Low")) %>%
  drop_na()

# box kernel
box_kernel2_reg <- ksmooth(x = df$Wind, y = df$Temp,
                          kernel = 'box', bandwidth = 2)
box_kernel5_reg <- ksmooth(x = df$Wind, y = df$Temp,
                          kernel = 'box', bandwidth = 5,
                          x.points = c(5, 10, 15, 20))
# gaussian normal kernel
gauss_kernel5_reg <- ksmooth(x = df$Wind, y = df$Temp,
                          kernel = 'normal', bandwidth = 3)

# this could be made in ggplot2, but I've skipped it for speed
#   following the help file instead
ggplot(df, aes(x = Wind, y = Temp)) + geom_point() +
  geom_line(data = as.data.frame(box_kernel2_reg), aes(x, y), color = 'red') +
  geom_line(data = as.data.frame(gauss_kernel5_reg), aes(x, y), color = 'blue')
```


### Locally Weighted Regression

```{r}
library(KernSmooth)

# local linear regression
df <- airquality %>% 
  mutate(high_temp = ifelse(Temp > 75, "High", "Low")) %>%
  drop_na()

air_loclin <- locpoly(x = df$Wind, y = df$Temp,
                      degree = 1, bandwidth = 5)

# local polynomial
air_locpoly <- locpoly(x = df$Wind, y = df$Temp,
                      degree = 2, bandwidth = 5)

# plotted
ggplot(df, aes(x = Wind, y = Temp)) + geom_point() +
  geom_line(data = as.data.frame(air_loclin), aes(x, y), color = 'red') +
  geom_line(data = as.data.frame(air_locpoly), aes(x, y), color = 'blue')
```


