---
title: "11_MLPS_R_ensemble_learning"
author: "Zhe Zhang (TA - Heinz CMU PhD)"
date: "3/07/2017"
output: 
  html_document:
    css: '~/Dropbox/avenir-white.css'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F, message = F)
```

## Lecture 11: Ensemble Learning

Key specific tasks we covered in this lecture:

* bootstrap aggregating (bagging)
* random forests (bagging, with selecting number of features at each node)
* adaptive boosting (adaBoost)

### Tree Ensembles

You can use the `randomForest` package for both bagging and boosting. They will differ depending on how they choose features for each node.

```{r}
library(tidyverse)
library(randomForest)

num_attributes = length(names(iris)) - 1
# bagging
bagged_iris <- randomForest(Species ~ ., data = iris,
                            type = 'classification',
                            mtry = num_attributes)

# forest
forest_iris <- randomForest(Species ~ ., data = iris,
                            type = 'classification',
                            mtry = sqrt(num_attributes))

# predictions (in-sample for brevity)
predict(forest_iris,
        newdata = iris,
        type = 'vote') %>% head(15)
predict(forest_iris,
        newdata = iris,
        type = 'response') %>% head(15)
predict(forest_iris,
        newdata = iris,
        type = 'prob') %>% head(15)
```


### adaBoost

The `ada` package builds on top of the `rpart` package to generate the weak learners.

```{r}
library(ada)

dat <- airquality %>% 
  mutate(high_temp = ifelse(Temp > 75, "High", "Low")) %>%
  drop_na() %>%
  select(-Month, -Day, -Temp) %>%
  mutate_if(.funs = scale, .predicate = is.numeric)

splits <- sample(c(0,1), size = nrow(dat), r = T, prob = c(2,1))
train <- dat[splits == 0, ]
test <- dat[splits == 1, ]

# enforcing decision stumps
#   
ada_air <- ada(high_temp ~ ., data = train,
               test.x = test %>% select(-high_temp),
               test.y = test$high_temp,
               type = 'discrete',
               rpart.control(maxdepth=1,cp=-1,minsplit=0,xval=0))

# get predictions
predict(ada_air, 
        newdata = test,
        type = 'vector') %>% head()
# getting probabilities are tricky,
#   both in interpretation
#   and in making sure you know which column represents what
predict(ada_air, 
        newdata = test,
        type = 'probs') %>% head()

# quick built in plot
plot(ada_air, test = T)
```

