---
title: "2_MLPS_R_visualization"
author: "Zhe Zhang (TA - Heinz CMU PhD)"
date: "1/19/2017"
output: 
  html_document:
    css: '~/Dropbox/avenir-white.css'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F)
```

## Lecture 2: Business & Data Understanding

Key specific tasks we covered in this lecture are around exploratory data analysis.
* dataset attribute types
* working with categorical variables (`factor` variables in R)
* data matrix styles (long vs wide)
* checking for data quality 

* temporal line plot
* bar chart
* histogram and 1-dimensional density visualization (and bin sizes)
* 2-dimensional histograms
* kernel density estimation
* median, quantile, quartile, interquartile range
* box plots
* scatter plots (many plots, overplotting, correlation graphs)
* parallel coordinates
* radar plots

## Dataset Actions

```{r}
# use the flights dataset
require(tidyverse)

# use `glimpse()` or `str()` to see a dataset's attributes
glimpse(iris)
glimpse(mtcars)
```

When your data has categorical values, if dataset has already been cleaned, you may not need to do much with it except be aware of it. When there are issues or things you'd like to change about your categorical variables, use the `forcats` package. Also, it can help reduce errors to not use factors in the first place, the `tidyverse` packages should work just as well with text/character variables versus factors. To convert a column to a factor use `{r} df$factor_column <- as.factor(df$factor_column)`.

To convert a dataset between a "wide" format and a "long" format, use the `gather` and `spread` functions available in the tidyr package. See more here: <http://r4ds.had.co.nz/tidy-data.html#tidy-data-1>

```{r}
library(tidyverse)

table4a

# renaming to avoid the weird columns that start with numbers
renamed_table4a <- table4a %>% rename('yr_1999' = `1999`, 'yr_2000' = `2000`)

# gather the data to make it long and make each row an observation
renamed_table4a %>% 
  gather(yr_1999, yr_2000, key = "year", value = "cases")
```

See here <http://r4ds.had.co.nz/tidy-data.html#spreading> for more information on making data from long to wide.

To check for data quality and potential missing variables, two keys options are summarization and visualization (and both combined). We discuss visualization below. This will not catch all potential errors, but is a way to start. This is especially useful when you have subgroups or categorical variables from which to compare. This chapter covers a lot of the EDA that we talked about in class in R: <http://r4ds.had.co.nz/exploratory-data-analysis.html>

```{r}
library(dplyr)
library(nycflights13)

by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))

by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")

# It looks like delays increase with distance up to ~750 miles 
# and then decrease. Maybe as flights get longer there's more 
# ability to make up delays in the air?
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

# cite: R4DS
```


## Visualization Actions

The R4DS has more examples and a thorough discussion of both visualization and especially, important, how to use visualization for exploratory analysis, which is focus of the early part of this course. I highly recommend reading through these two sections and looking at the examples. If you're not already comfortable with the pieces, doing so will be quite helpful for all your work in this class and going forward with R and analysis. I will only briefly touch on the specific commands below.

* <http://r4ds.had.co.nz/data-visualisation.html>
* <http://r4ds.had.co.nz/exploratory-data-analysis.html>

```{r}
# More specific examples of commands incoming
#   mostly based around ggplot2 package
library(tidyverse)
library(ggplot2) # redundant, but here for emphasis

# temporal line plot
# bar chart
# histogram and 1-dimensional density visualization (and bin sizes)
# 2-dimensional histograms
# kernel density estimation
# median, quantile, quartile, interquartile range
# box plots
# scatter plots (many plots, overplotting, correlation graphs)
# parallel coordinates
# radar plots
```

This is incomplete at the moment, apologies. I will make an announcment on Piazza once this is complete, likely by Friday morning.
