---
title: "9_MLPS_R_SVM_Kernels"
author: "Zhe Zhang (TA - Heinz CMU PhD)"
date: "3/04/2017"
output: 
  html_document:
    css: '~/Dropbox/avenir-white.css'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F, message = F)
```

## Lecture 9: Support Vector Machines & Kernels

Key specific tasks we covered in this lecture:

* basic SVM fit
* SVM with hard margin
* normalizing input variables for SVM
* RBF, polynomial, sigmoid kernel feature transformation
* SVM regularization parameter C
* kernel PCA


### Basic SVM Examples

There are several possible SVM packages, we will discuss `e1071`.

```{r}
library(tidyverse)
library(e1071)

df <- airquality %>% 
  mutate(high_temp = ifelse(Temp > 75, 1, 0)) %>%
  drop_na()

wrong_svm <- svm(high_temp ~ Ozone + Solar.R + Wind, data = df)
summary(wrong_svm)
```

Be careful about ensuring that you're performing classification.

```{r}
# to get the correct classification SVM: 
#   either change the outcome to a factor
#   or specify explicitly
svm_classifier <- svm(high_temp ~ Ozone + Solar.R + Wind, 
                      data = df,
                      type = 'C-classification')

df <- df %>% mutate(high_temp = as.factor(high_temp))
svm_classifier <- svm(high_temp ~ Ozone + Solar.R + Wind, 
                      data = df)
summary(svm_classifier)
```

### Adjusting SVM options (`C`, normalizing features, kernel trick transformation)

```{r}
# we specify normalizing the (numeric) features
#   the kernel to use
#   the cost (C) parameter for the soft margin

svm_linear_kernel <- 
  svm(high_temp ~ Ozone + Solar.R + Wind, 
      data = df, kernel = 'linear', scale = T,
      cost = 0.5,
      type = 'C-classification')

svm_polynomial_kernel <- 
  svm(high_temp ~ Ozone + Solar.R + Wind, 
      data = df, kernel = 'polynomial', scale = T,
      cost = 0.75, degree = 2, gamma = 0.1,
      type = 'C-classification')

svm_rbf_kernel <- 
  svm(high_temp ~ Ozone + Solar.R + Wind, 
      data = df, kernel = 'radial', scale = T,
      cost = 0.75, gamma = 0.25,
      type = 'C-classification')

# for sigmoid kernel, use kernel = 'sigmoid`

# to get probabilities, use the probability option
svm_rbf_kernel <- 
  svm(high_temp ~ Ozone + Solar.R + Wind, 
      data = df, kernel = 'radial', scale = T,
      cost = 0.75, gamma = 0.25,
      probability = T,
      type = 'C-classification')
```

### Predictions and Plotting SVM Fits

```{r}
# predictions
svm_rbf_preds <- predict(svm_rbf_kernel, newdata = df)

svm_rbf_probability <- predict(svm_rbf_kernel, newdata = df,
                               probability = T) %>%
  attr("probabilities") %>% as_data_frame()

# simple 2-D plots
#   the predicted classification region is color
#   the actual ground truth is in black "X" or red "O"
plot(svm_rbf_kernel, 
     data = df, 
     formula = Ozone ~ Wind) # which dimensions to compare
plot(svm_linear_kernel, 
     data = df, 
     formula = Ozone ~ Wind)
```


# Kernel PCA

```{r}
library(kernlab)
# another example using the iris
data(iris)
test <- sample(1:150,20)

kpc <- kpca(~.,data=iris[-test,-5],kernel="rbfdot",
            kpar=list(sigma=0.2),features=2)

#print the principal component vectors
pcv(kpc) %>% head(25)

#plot the data projection on the components
plot(rotated(kpc),col=as.integer(iris[-test,5]),
     xlab="1st Principal Component",ylab="2nd Principal Component")
```

